Exploring Molecular Conformations and Related Time Scales via Nonequilibrium Molecular Dynamics Simulation

### 2013-08-06 writedown/npt-neq/alanine.tex
back up, first finished version

The possible effect of the electric field (EF) on the conformation of
proteins gains increasing attantion recently, because it closly
related to the health problem of exposing human tissue in the
electromagnetic radiation. Despite the importance of this problem, a
complete understanding of how the electric field influnece the  protein
is still lacking, therefore, a conserderable number of
experimental~\cite{bohr2000microwave, bohr2000microwave-1,
  dePomerai2000cell, inskip2001cellular, mancinelli2004non} and
theoretical~\cite{budi2005electric, budi2007effect,
  budi2008comparative, toschi2008effects, astrakas2011electric,
  astrakas2012structural, damm2012can, starzyk2013proteins,
  english2009nonequilibrium, solomentsev2012effects} studies have been devoted to this topic,
and the references listed above is far from complete.

The molecular dynamics (MD) simulation has been proved to be useful in
understanding the behavior of a protein (or a short peptide segment)
in both static and oscillatory electiric
fields~\cite{budi2005electric, budi2007effect, budi2008comparative,
  toschi2008effects, astrakas2011electric, astrakas2012structural,
  damm2012can, starzyk2013proteins, english2009nonequilibrium,
  solomentsev2012effects}. In these studies, the initial
configurations are usually taken from the Protein Data Bank (PDB).
And then the temperature controlled simulation are performed for from
a few to tens of nanoseconds, during which the interested
conformational chanages can be observed. Along the trajectories, the
secondary structure, root mean square displacement (RMSD), the dipole
moment and some shape parameters are calculated and investigated by
either plotted as a function of time, or doing time average over the
whole simulation. This approach release useful information on how the
secondary structure is changed by external electiric field.  However,
by plotting the quantities against time, one may observe strong
fluctuation in plots, which hintered the quantitative and conclusive
analysis on what is happening during the conformational change, when
it happens and what is the strength. Although the statistical
uncertainty in the time-averaged observations are small, the
time-average itself is not well defined for a nonequilibrium process
such as relaxing process to the new conformations under static
electiric field, and the dynamics under oscillating electric field.

The dynamical nonequilibrium molecular dyanmics simulation (D-NEMD)
was firstly developed by G. Ciccotti and his
colleagues~\cite{ciccotti1975direct, ciccotti1979thought}, and
recently successfully applied to the study of
hydrodynamics~\cite{orlandini2011hydrodynamics,
  orlandini2011hydrodynamics-01}. D-NEMD provides a way of evaluating
nonequilibrium observables by doing an ensemble of nonequilibrium MD
simualtions.  This provides a promising way of analysing the
conformational changing under a generally time-dependent electric
field. In the present paper, we study the minimal model peptide
segment -- alanine dipeptide. We take the approach of starting from a
single peptide rather than a more realistic chain, firstly because it
is easier to draw clear conclusions on the nonequilibrium properties
of a single peptide by excluding all possible interplay between
different peptide segments along a chain, then construct the chain
nonequilibrium properties from the well-studied single peptides.
Secondly, the current computationl resource available is not enought
to study peptide chains by D-NEMD, since one needs to simualte an
ensemble of MD trajectories (typically a few thousands), so it
requires much higher computational cost than producing a single MD
trajectory.

In this paper, we firstly describe the idea of the D-NEMD, then
observables helping us the analyze the nonequilibrium properties
(i.e., the dipole moment, time-dependent properties of being a certain
conformation, the nonequilibrium transition properties and the net
property flux) are introduced. The numerical results of alanine
dipeptide under static and sine-shaped oscillating electric field are
presented. By calculating the nonequilibrium properties from the D-NEMD
simulations, we are able to answer from which to which the conformation
changes, when do these changes happend, and how strong are the changes.
As far as we know, our simulation is the first D-NEMD simulation
on the electric field effect on alanine dipeptide, and provides new
information helping people better understand the nonequilibrium
effect of electric field on the protein conformation.

\section{Methodology}
\subsection{Nonequilibrium molecular dynamics simulation}
Here we fastly review the approach of performing nonequilibrium
molecular dynamics that was initiated by Giovanni Ciccoti and his
colleagues~\cite{ciccotti1975direct, ciccotti1979thought,
  orlandini2011hydrodynamics, orlandini2011hydrodynamics-01}.
We denote the macroscopic nonequilibrium observable by $O(t)$. If at $t$
the configurational probability distribution is $\rho(\vect x, t)$, where
$\vect x$ is the phase space variable, then the observable can be
expressed by
\begin{align}\label{eqn:tmp1}
  O(t) = \int d\vect x\, \hat O(\vect x)\rho(\vect x, t)  = \langle \hat O(\vect x), \rho(\vect x, t)\rangle,
\end{align}
where $\hat O (\vect x)$ is the microscopic observable. Compared with
the macroscopic observable, it is measured at the phase space position $\vect x$.
The bracket on the right hand side denotes the inner produce in the
phase space.  We assume the dynamics of the system is governed by the
Hamiltonian equation, i.e. $\dot {\vect x} = J \cdot \nabla_{\vect x}
\mathcal H(\vect x)$, where $\mathcal H$ is the Hamiltonian and $J$ is
the symplectic matrix, then the Liouville equation of the probability
distribution is
\begin{align}\label{eqn:tmp2}
  \frac{\partial \rho(\vect x, t)}{\partial t} = - iL(t) \rho(\vect x, t),
\end{align}
where $iL(t) = \{\cdot, \mathcal H\}$ is the Liouville operator and
$\mathcal H$ is the Hamiltonian of the system. The Eq.~\eqref{eqn:tmp2}
can be formally solved by $\rho(\vect x, t) = e^{-iL(t)} \rho(\vect x, 0)$.
On the other hand
\begin{align}
  \frac{d \hat O(\vect x(t))}{dt} = \nabla_{\vect x}\hat O\cdot \dot{\vect x}
  = \nabla_{\vect x}\hat O\cdot J\cdot \nabla_{\vect x}\mathcal H
  = iL(t) \hat O (\vect x(t))
\end{align}
This equation can be formally solved by $\hat O(\vect x(t)) = e^{iL(t)} O(\vect x, 0)$, therefore,
\begin{align}\nonumber
  O(t) & = \langle \hat O(\vect x), \rho(\vect x, t)\rangle
  = \langle \hat O(\vect x), e^{-iL(t)} \rho(\vect x, 0)\rangle
  = \langle e^{iL(t)}\hat O(\vect x), \rho(\vect x, 0)\rangle\\\label{eqn:tmp4}
  &= \langle \hat O(\vect x(t)), \rho(\vect x, 0)\rangle
\end{align}
If the inital state of the system is in equilibrium, i.e. the initial
probability distribution is the equilibrium distribution. Then an
explicit translation of Eq.~\eqref{eqn:tmp4} is that the
nonequilibrium observable is equal to the ensenble average of
microscopic observable measured along trajectory $\vect x(t)$, the
inital configuration of which is subject to the equilibrium
distribution. In practice, we firstly run an equilibrium MD simulation
to generate a collection of configurations that are subject to the
equilibrium distribution. And then by using these configurations as
inital configuration, the Hamiltonian dynamicses are integrated until
time $t$. These trajectories are also called \emph{branching
  trajectories} in this paper. Finially, the macroscopic observable is
estimated by averaging the microscopic observable measured at the end
points of the trajectories.

Being more gneralize, if the microscopic depends not only on the end
configuation of at time $t$ but also depends on the historical
configurations along the trajectory, the experssion~\eqref{eqn:tmp4} can be generalize to
\begin{align}
  O(t) = \int d\vect x\,\rho(\vect x, 0) \int_{\mathcal C\{\vect x, 0; t\}} \hat O[\vect x_s] \,d\mathcal P[\vect x_s] 
\end{align}
where the microscopic observable $ \hat O[\vect x_s] $ is now a
functional of the trajectory.  $\mathcal C\{\vect x, 0; t\}$ is the
set of all continuous trajectories starting at point $\vect x$ at time
0, and ending at time $t$. $\mathcal P[\vect x_s] $ is the meansure of
the trajectory space $\mathcal C$.  In our case, it is the
$\delta$-function peaked around the trajectory generated by the
Hamiltonian dynamics. To be general, it can also be the measure of all
trajectories generated by, for example, a Langevin dynamics, then the
branching trajectories should be generated by the Langevin
dynamics rather than the Hamiltonian dynamics.

@@

The aformentioned nonequilibrium molecular dynamicses simulation algorithm
does not describe how the boundary of the system should be set up. The 100\% save
method is to simulate a infinitely large isolated system, or any large enough
isolated subsystem of it. In practice, it is both impossible and unnecessary to
do such a simulation. In practice, proper boundary condition should be developped
based on how the nonequilibrium system is set up, and what are the nonequilibrium
observables to measure.
Since we are mainly interested in the nonthermal~\cite{delaHoz2005microwaves} effect of
the electric field to the configuration of a solvated alanine dipeptide,
we assume the alanine is embeded into an infinitely large solvent envrionment,
and the electric field is only applied to the neighborhood of the alanine.
Therefore, 
extra heat generated by turing on the electric field can be effectively absorbed
by the envrionment.
It is argued that 
an accurate reaction temperature control is essential for being able to perform
reproducible experiments~\cite{damm2012can}.

@@

In practice, it is impossible to simulate an infinitely large
environment.  For an equilibrium simulation, the solution is to couple
the system to a thermostat, under which the canonical ensember can be
sampled by only simulate a finite size system with periodic boundary
condition. For a nonequilibrium simulation, simply couple the system
to a thermostat may not be a good idea, because the \emph{dynamics}
generated by the thermostat is artificial, and deviation from the
Hamiltonian dynamics under investigation will introduce artificial
effects in the nonequilibrium observations. To solve this problem, we
observe that only the Hamiltonian dynamics of the alanine molecule
itself and the water molecules in the nearby solvation shell is
crutial to the configuration change, while the detailed dynamics of
the water molecules far away has less effect. Notice again we actually
do not want to precisely calculate each trajectory starting from the
initial configurations, but want to sample correctly the trajectory
ensemble correctly, so that the nonequilibrium observable can be
calculated correctly.  Therefore, instead of simulating an infinitely
large system, we only simulate a finite size system with periodic
boundary condition. The simulating region is divided into two
subregions: Near the alanine dipeptide, the dynamics is kept to be
Hamiltonian, and this subregion is called the Hamiltonian dynamics
region. While far from the alanine, the dynamics of water is coupled
to a Langevin thermostat, and this region is called the thermostating
region (see Fig.~\ref{fig:tmp2}).  For simplicity, the Hamiltonian
dynamics region is assumed to be spherical, centered at the
alpha-carbon of the alanine, and with radius of $R_{ex}$.  Further, we
freeze the movment of the alpha-carbon, so that the alanine is always
located at the center of the simulation region.  The validity of this
setting will be later check by numerical examples showing that the
nonequilibrium observables do not depends on the size of the system
and the size of the Hamiltonian dynamics subregion, if they are
reasonably large and the finite size effect does not play a role.

@@

We want to study the nonequilibrium properties of an alanine dipeptide
under a changing electric field. As a model system, we study the
a constant electric field. At time $t=0$~ps, the system
has been fully equilibriated without any electric field. From $t=0$ to
$t=t_{\textrm{init}}$, the electic field is switched on linearly, while
after $t=t_{\textrm{init}}$, the field is kept constant at
$E_{\infty}$. Fig.~\ref{fig:tmp3} shows the three different
cases considered in this work: $t_{\textrm{init}} = 10$, 100, and
500~ps, and 
in all cases the maximum strength of the electric field is
$E_{\infty} = 1$~V/nm.
The direction of this field is arbitarily chosen: along the
$x$ direction. Since the dipole of the equilibrium configurations of
alanine can be along any direction (see also Fig.~\ref{fig:tmp3} for
the vanished initial dipole moment), applying the electric along x-axis is
not biased.
The size of the system is $2.67\times 2.67\times
2.67\, \textsf{nm}^3$, with one alanine dipeptide described by the CHARMM27 force field, and solved in 644 TIP3P
water molecules.
All simulations are performed by a home-modified GROMACS 4.5~\cite{pronk2013gromacs}.
Firstly, an equilibrium NVT simulation of
100~\textsf{nm} was performed with a Langevin thermostat (time-scale
$\tau_T = 0.5~\textsf{ps}$) coupled to the system for constant
temperature.  Along the trajectory, configurations were taken every
5~\textsf{ps}.  They server as starting point of the branching
trajectories. The branching trajectories were integrated by the
Leap-frog scheme with the aformentioned nonequilibrium temperature control technique.  The
time step was $\Delta t = 0.002~\textsf{ps}$. The short-range
interaction was smoothed from $0.8$ to $1.0~\textsf{nm}$ by the
``switch'' method provided by Gromacs.  An energy conserving PME
method was applied to calculate the electrostatic interaction in this
periodic system. In the thermostating region, the original dyanmics was
coupled to a Langevin thermostat with $\tau_T = 0.1~\textsf{ps}$.
The whole system is also coupled to a Parrinello-Rahman barostat with $\tau_P = 2.0~\textsf{ps}$ to keep
the pressure at the ambient condition (1~Bar). Since the
change of the system size is small and slow, it does not have an obviously effect on the
dyanmics of the system.

@@

On each branching trajectory, to easy the analysis, the conformation
of the alanine dipeptide is plotted by the Ramachandran histogram, as
shown in Fig.~\ref{fig:tmp4}. The equilibrium distribution of the
conformation is given in the left plot, and the fully relaxed
conformation at $E_{\infty} = 1$~V/nm is given in the right plot.  It is
clear the conformations are populated in several clusters, and the
position of the clusters do not change when the external electric
field is applied. This means that only a few of the configurations are likely to be observed, while there is nearly
no oppotunity to see others. 
These configurations are called ``metastable states''. The strict definition of the
metastability is not exactly the same as those conformations that are likely to be observed.
Since in this paper we do not focus on the discussion of the definition for the metastability of the system,
we assume they have the same meaning.
We manually divide the
$\psi$--$\phi$ plane into 5 subregions, see Fig.~\ref{fig:tmp4},
then the observation on the Ramachandran histogram is projected to the
observation of these 5 subregions.  These subregions
correspond to different secondary structure of a peptide chain:
$A_1$ and $A_2$ are
corresponding to the alpha helix. $B_1$ and $B_2$ are
corresponding to beta sheet. $C$ is corresponding to the
left-handed alpha helix.

\begin{figure}
  \centering
  \includegraphics[width=0.45\textwidth]{fig.ala/ext.mode1.100.Ex.01.00.t1200ps/fig-begin-1.eps}
  \includegraphics[width=0.45\textwidth]{fig.ala/ext.mode1.100.Ex.01.00.t1200ps/fig-end-1.eps}
  \caption{The population of different metastable state. Left: the equilibrium state at $t=0$~\textsf{ps}. Right: The non-equilibrium study state at $t=1200$~\textsf{ps}, i.e., fully relaxed when the electric field is switched on. The darker color indicates larger population on the $\psi$--$\phi$ phase space.}
  \label{fig:tmp4}
\end{figure}

>>>>

The possible effect of the electric field (EF) on the conformation of
proteins gains increasing attention recently, because it closely
related to the health problem of exposing human tissue in the
electromagnetic radiation. Despite the importance of this problem, a
complete understanding of how the electric field influence the  protein
is still lacking, therefore, a considerable number of
experimental~\cite{bohr2000microwave, bohr2000microwave-1,
  dePomerai2000cell, inskip2001cellular, mancinelli2004non} and
theoretical~\cite{budi2005electric, budi2007effect,
  budi2008comparative, toschi2008effects, astrakas2011electric,
  astrakas2012structural, damm2012can, starzyk2013proteins,
  english2009nonequilibrium, solomentsev2012effects} studies have been devoted to this topic,
and the references listed above is far from complete.

The molecular dynamics (MD) simulation has been proved to be useful in
understanding the behavior of a protein (or a short peptide segment)
in both static and oscillatory electric
fields~\cite{budi2005electric, budi2007effect, budi2008comparative,
  toschi2008effects, astrakas2011electric, astrakas2012structural,
  damm2012can, starzyk2013proteins, english2009nonequilibrium,
  solomentsev2012effects}. In these studies, the initial
configurations are usually taken from the Protein Data Bank (PDB).
And then the temperature controlled simulation are performed for from
a few to tens of nanoseconds, during which the interested
conformational changes can be observed. Along the trajectories, the
secondary structure, root mean square displacement (RMSD), the dipole
moment and some shape parameters are calculated and investigated by
either plotted as a function of time, or doing time average over the
whole simulation. This approach release useful information on how the
secondary structure is changed by external electric field.  However,
by plotting the quantities against time, one may observe strong
fluctuation in plots, which hintered the quantitative and conclusive
analysis on what is happening during the conformational change, when
it happens and what is the strength. Although the statistical
uncertainty in the time-averaged observations are small, the
time-average itself is not well defined for a nonequilibrium process
such as relaxing process to the new conformations under static
electric field, and the dynamics under oscillating electric field.

The dynamical nonequilibrium molecular dynamics simulation (D-NEMD)
was firstly developed by G. Ciccotti and his
colleagues~\cite{ciccotti1975direct, ciccotti1979thought}, and
recently successfully applied to the study of
hydrodynamics~\cite{orlandini2011hydrodynamics,
  orlandini2011hydrodynamics-01}. D-NEMD provides a way of evaluating
nonequilibrium observables by doing an ensemble of nonequilibrium MD
simulations.  This provides a promising way of analyzing the
conformational changing under a generally time-dependent electric
field. In the present paper, we study the minimal model peptide
segment -- alanine dipeptide. We take the approach of starting from a
single peptide rather than a more realistic chain, firstly because it
is easier to draw clear conclusions on the nonequilibrium properties
of a single peptide by excluding all possible interplay between
different peptide segments along a chain, then construct the chain
nonequilibrium properties from the well-studied single peptides.
Secondly, the current computational resource available is not enough
to study peptide chains by D-NEMD, since one needs to simulate an
ensemble of MD trajectories (typically a few thousands), so it
requires much higher computational cost than producing a single MD
trajectory.

In this paper, we firstly describe the idea of the D-NEMD, then
observables helping us the analyze the nonequilibrium properties
(i.e., the dipole moment, time-dependent properties of being a certain
conformation, the nonequilibrium transition properties and the net
property flux) are introduced. The numerical results of alanine
dipeptide under static and sine-shaped oscillating electric field are
presented. By calculating the nonequilibrium properties from the D-NEMD
simulations, we are able to answer from which to which the conformation
changes, when do these changes happens, and how strong are the changes.
As far as we know, our simulation is the first D-NEMD simulation
on the electric field effect on alanine dipeptide, and provides new
information helping people better understand the nonequilibrium
effect of electric field on the protein conformation.

\section{Methodology}
\subsection{Nonequilibrium molecular dynamics simulation}
Here we fastly review the approach of performing nonequilibrium
molecular dynamics that was initiated by Giovanni Ciccoti and his
colleagues~\cite{ciccotti1975direct, ciccotti1979thought,
  orlandini2011hydrodynamics, orlandini2011hydrodynamics-01}.
We denote the macroscopic nonequilibrium observable by $O(t)$. If at $t$
the configurational probability distribution is $\rho(\vect x, t)$, where
$\vect x$ is the phase space variable, then the observable can be
expressed by
\begin{align}\label{eqn:tmp1}
  O(t) = \int d\vect x\, \hat O(\vect x)\rho(\vect x, t)  = \langle \hat O(\vect x), \rho(\vect x, t)\rangle,
\end{align}
where $\hat O (\vect x)$ is the microscopic observable, 
which is measured at the phase space position $\vect x$,
comparing with the macroscopic observable that is not phase space independent.
In this work, we always assume that the initial probability distribution
$\rho(\vect x, 0)$ is known, and is equal to the equilibrium distribution
without any electric field.
The bracket on the right hand side of Eq.~\eqref{eqn:tmp1} denotes the inner produce in the
phase space.  We assume the dynamics of the system is governed by the
Hamiltonian equation, i.e. $\dot {\vect x} = J \cdot \nabla_{\vect x}
\mathcal H(\vect x)$, where $\mathcal H$ is the Hamiltonian, and $J$ is
the symplectic matrix, then the Liouville equation of the probability
distribution is
\begin{align}\label{eqn:tmp2}
  \frac{\partial \rho(\vect x, t)}{\partial t} = - iL(t) \rho(\vect x, t),
\end{align}
where $iL(t) = \{\cdot, \mathcal H\}$ is the Liouville operator.
The Eq.~\eqref{eqn:tmp2}
can be formally solved by $\rho(\vect x, t) = e^{-iL(t)} \rho(\vect x, 0)$.
On the other hand
\begin{align}
  \frac{d \hat O(\vect x(t))}{dt} = \nabla_{\vect x}\hat O\cdot \dot{\vect x}
  = \nabla_{\vect x}\hat O\cdot J\cdot \nabla_{\vect x}\mathcal H
  = iL(t) \hat O (\vect x(t))
\end{align}
This equation can be formally solved by $\hat O(\vect x(t)) = e^{iL(t)} O(\vect x, 0)$, therefore,
\begin{align}\nonumber
  O(t) & = \langle \hat O(\vect x), \rho(\vect x, t)\rangle
  = \langle \hat O(\vect x), e^{-iL(t)} \rho(\vect x, 0)\rangle
  = \langle e^{iL(t)}\hat O(\vect x), \rho(\vect x, 0)\rangle\\\label{eqn:tmp4}
  &= \langle \hat O(\vect x(t)), \rho(\vect x, 0)\rangle
\end{align}
Since we assume that the initial
probability distribution is the equilibrium distribution, an
explicit translation of Eq.~\eqref{eqn:tmp4} is that the
nonequilibrium observable is equal to the ensemble average of
microscopic observable measured along trajectory $\vect x(t)$, the
initial configuration of which is subject to the \emph{equilibrium}
distribution. In practice, we firstly run an equilibrium MD simulation
to generate a sampling of configurations that are subject to the
equilibrium distribution. And then by using these configurations as
initial configuration, the Hamiltonian dynamics are integrated until
time $t$. These trajectories are also called \emph{branching
  trajectories} in this paper. Finally, the macroscopic observable is
estimated by averaging the microscopic observable measured at the end
points of the trajectories.

Being more generalized, if the microscopic observable depends not only on the end
configuration at time $t$ but also depends on the historical
configurations along the trajectory, the expression~\eqref{eqn:tmp4} can be generalize to
\begin{align}
  O(t) = \int d\vect x\,\rho(\vect x, 0) \int_{\mathcal C\{\vect x, 0; t\}} \hat O[\vect x_s] \,d\mathcal P[\vect x_s] 
\end{align}
where the microscopic observable $ \hat O[\vect x_s] $ is now a
functional of the trajectory.  $\mathcal C\{\vect x, 0; t\}$ is the
set of all continuous trajectories starting at point $\vect x$ at time
0, and ending at time $t$. $\mathcal P[\vect x_s] $ is the measure of
the trajectory space $\mathcal C$.  In our case, it is the
$\delta$-function peaked around the trajectories governed by the
Hamiltonian dynamics. To be general, it can also be the measure of all
trajectories generated by, for example, a Langevin dynamics, then the
branching trajectories should be generated by the Langevin
dynamics rather than the Hamiltonian dynamics.

@@

The aforementioned nonequilibrium molecular dynamics simulation algorithm
does not describe how the boundary of the system should be set up.
The very naive idea of doing the simulation
is to simulate a infinitely large isolated system, or any large enough
isolated subsystem of it. In practice, it is both impossible and unnecessary to
do such a simulation, due to the unaffordable computational expense.
Therefore, proper boundary condition should be supplied
to truncated the simulation region. It is only possible when we
have a clear knowledge on
on how the nonequilibrium system is set up,
and what are the nonequilibrium observables to measure.

@@

Since we are mainly interested in the
nonthermal~\cite{delaHoz2005microwaves} effect of the electric field
to the configuration of a solvated alanine dipeptide, we assume the
alanine is embeded into an infinitely large solvent environment, and
the electric field is only applied to the neighborhood of the alanine.
Any extra heat generated by the electric field can be effectively
absorbed by the solvent environment.  It is argued that such an
accurate reaction temperature control is essential for being able to
perform reproducible experiments~\cite{damm2012can}.
Now the question comes to how to truncate the solvent environment
so that the system can be handled by the limited computer resource.
% In practice, it is impossible to simulate an infinitely large
% environment.
For an equilibrium simulation, the solution is to couple
the system to a thermostat, under which the canonical ensemble can be
sampled by only simulating a finite size system with periodic boundary
condition. For a nonequilibrium simulation, simply couple the system
to a thermostat may not be a good idea, because the \emph{dynamics}
generated by the thermostat is artificial, and deviation from the
Hamiltonian dynamics under investigation will introduce artificial
effects in the nonequilibrium observations. To solve this problem, we
observe that only preserving the Hamiltonian dynamics of the alanine 
itself and the water molecules in the nearby solvation shell is
crucial to the interested nonequilibrium observable
(in this work, it is the conformational change of alanine),
while the detailed dynamics of
the water molecules far away is of less importance. 
Therefore, instead of simulating an infinitely
large system, we only simulate a finite size system with periodic
boundary condition. The simulating region is divided into two
subregions: Near the alanine dipeptide, the dynamics is kept to be
Hamiltonian, and this subregion is called \emph{the Hamiltonian dynamics
region}. While far from the alanine, the dynamics of water is coupled
to a Langevin thermostat, and this region is called \emph{the thermostating
region} (see Fig.~\ref{fig:tmp2}).  For simplicity, the Hamiltonian
dynamics region is assumed to be spherical, centered at the
alpha-carbon of the alanine, and with radius of $R_{ex}$.  Further, we
freeze the movement of the alpha-carbon, so that the alanine is always
located at the center of the simulation region.  The validity of this
setting will be later check by numerical examples showing that the
nonequilibrium observables do not depends on the size of the system
and the size of the Hamiltonian dynamics subregion, if they are
reasonably large and the finite size effect does not play a dominating role.

% Notice again we actually
% do not want to precisely calculate each trajectory starting from the
% initial configurations,
% but want to correctly calculate the nonequilibrium observable,
% we

@@

In this section,
we want to study the nonequilibrium properties of an alanine dipeptide
under a constant EF. At time $t=0$~ps, the system
has been fully equilibriated without any electric field. From $t=0$ to
$t=t_{\textrm{init}}$, the eclectic field is switched on linearly, while
after $t=t_{\textrm{init}}$, the field is kept constant at
$E_{\infty}$. In Fig.~\ref{fig:tmp3}, the dashed line shows the 
cases considered in this work: $t_{\textrm{init}} = 10$~ps,
and $E_{\infty} = 1$~V/nm.
The direction of this field is arbitrarily chosen: along the
$x$ direction. Since the dipole of the equilibrium configurations of
alanine can be along any direction (see also Fig.~\ref{fig:tmp3} for
the vanished initial dipole moment), applying the electric along $x$-axis is
not biased.
The system is set up in a $2.7\times 2.7\times
2.7\, \textsf{nm}^3$ periodic simulation region, with one alanine dipeptide
described by the CHARMM27 force field, and solved in 644 TIP3P
water molecules.
The size of the  Hamiltonian dynamics region is $R_{ex} = 1.0$~nm.
All simulations are performed by a home-modified GROMACS 4.5~\cite{pronk2013gromacs}.
Firstly, an equilibrium NVT simulation of
100~\textsf{ns} was performed with a Langevin thermostat (time-scale
$\tau_T = 0.5~\textsf{ps}$) coupled to the system for constant
temperature.  Along the trajectory, configurations were taken every
50~\textsf{ps}, so we use 2000 initial configurations for each nonequilibrium
MD simulation, if not stated otherwise.
% They server as starting point of the branching
% trajectories.
The branching trajectories were integrated by the
Leap-frog scheme with the aforementioned nonequilibrium
temperature control technique.  The
time step was $\Delta t = 0.002~\textsf{ps}$. The short-range
interaction was smoothed from $0.8$ to $1.0~\textsf{nm}$ by the
``switch'' method provided by Gromacs.  An energy conserving PME
method was applied to calculate the electrostatic interaction in this
periodic system. In the thermostating region, the original dynamics was
coupled to a Langevin thermostat with $\tau_T = 0.1~\textsf{ps}$.
The whole system is also coupled to a Parrinello-Rahman barostat with $\tau_P = 2.0~\textsf{ps}$ to keep
the pressure at the ambient condition (1~Bar). Since the
change of the system size is small and slow, the pressure control
does not have an obviously effect on the
dynamics of the system.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.45\textwidth]{fig.ala/ext.mode1.100.Ex.01.00.t1200ps/fig-begin-1.eps}
  \includegraphics[width=0.45\textwidth]{fig.ala/ext.mode1.100.Ex.01.00.t1200ps/fig-end-1.eps}
  \caption{The population of metastable states of the alanine dipeptide.
    Left: the
    equilibrium state at $t=0$~\textsf{ps}. Right: The new equilibrium
    state at $t=1000$~\textsf{ps}, i.e., fully relaxed under the
    constant electric field 1~V/nm. The darker color
    indicates larger population on the Ramachandran histogram.}
  \label{fig:tmp4}
\end{figure}

On each branching trajectory, to easy the analysis of the conformation
of alanine dipeptide, the molecular configuration is projected on
the Ramachandran histogram, as
shown in Fig.~\ref{fig:tmp4}. The equilibrium distribution of the
conformation is given in the left plot, and the fully relaxed
conformation under $E_{\infty} = 1$~V/nm is given in the right plot.  It is
clear the conformations are populated in several clusters, and the
position of the clusters do not change when the external electric
field is applied.
This means that only a few of the conformations are likely to be observed,
while there is nearly no opportunity to see others. 
These conformations are called ``metastable states''.
The strict definition of the
metastability is not exactly the same as those conformations that are likely to be observed.
Since in this paper we do not focus on the discussion of the definition for the metastability of the system,
we assume they have the same meaning.
We manually divide the
Ramachandran histogram into 5 subregions, see Fig.~\ref{fig:tmp4},
then the observation on the Ramachandran histogram is projected to the
observation of these 5 states, i.e. $\{A_1, A_2, B_1, B_2, C\}$, which
correspond to the molecular conformations we are investigating in this work.
These subregions
correspond to different secondary structure of a peptide chain:
$A_1$ and $A_2$ are
corresponding to the alpha helix. $B_1$ and $B_2$ are
corresponding to the beta sheet. $C$ is corresponding to the
left-handed alpha helix.

From Fig.~\ref{fig:tmp4}, it is obvious that the population in each
metastable state changes due to the external electric field, for
example, the population of state $C$ vanishes in equilibrium,
but plays an important role when the electric field is applied. We
therefore want to study how the alanine changes from one conformation
to the other, how fast is the change, and the relations to the 
the electric field. 
Firstly, we calculate the probability of being in a certain state that
is defined by:
\begin{align}
  P_I(t) = \mathbb P (\vect x_t \in I), \quad  I \in \{A_1, A_2, B_1, B_2, C\},
\end{align}
please notice that the time dependency of the probability, so it is
a nonequilibrium observable.
% which is the time dependent probability of the alanine being in metastable
% state $I$.
% Secondly, we study the
% forward and backward transition probability of the metastable conformations, which is defined by
% \begin{align}
%   P^{\fwd}_{J,I}(t) & = \mathbb P( \vect x_{t+\Delta t} \in J | \vect x_t \in I) \\
%   P^{\bwd}_{J,I}(t) & = \mathbb P( \vect x_{t-\Delta t} \in J | \vect x_t \in I) 
%   % \qquad I, J \in \{A_1, A_2, B_1, B_2, C\},
% \end{align}
% where $\Delta t$ is the time-scale of the observation.
% For a revserible and homogeneous Markov process, it is easy to show
% that $P^{\fwd}_{J,I} = P^{\bwd}_{J,I}$, and both of them are time independent. For a general nonequilibrium
% process considered by the present work, this relation is genarlly not guarenteed.
Secondly,
we study the net probability flux from metastable state $I$ to $J$, defined by:
\begin{align}\nonumber
  F_{J,I}(t) & =\lim_{\Delta t\rightarrow 0} \frac1{\Delta t} [\mathbb P( \vect x_{t-\Delta t} \in J, \vect x_t \in I) - \mathbb P( \vect x_{t-\Delta t} \in I, \vect x_t \in J)], \\\label{eqn:tmp7}
  & J,I \in \{A_1, A_2, B_1, B_2, C\}.
\end{align}
which presents the rate of how much conformation $J$ is changed
into $I$. A positive value indicates the
net flux from $J$ to $I$,
while a negative value indicates a net flux from $I$ to $J$.
Now project the original nonequilibrium process $\vect x_t$
onto the discretized states $\{A_1, A_2, B_1, B_2, C\}$,
if the resulting reduced process 
is time-reversible, then it is obvious that
$ F_{J,I}(t) = 0$. In an other word, the quantity $ F_{J,I}(t) $ represents
the irreversibility of the reduced process on
$\{A_1, A_2, B_1, B_2, C\}$.
In practice, to measure $ F_{J,I}(t)$, one need firstly approximate the limit
in Eq.~\eqref{eqn:tmp7} by the finite difference,
then calculate the joint probabilities on
the right-hand-side from the molecular trajectories.
If one wants a precise finite difference approximation, a smaller $\Delta t$ is
preferable,
but then the statistical uncertainty of measuring the joint probabilities
is higher. In practice,
we use a $\Delta t$ that is small enough for a precise finite difference approximation,
and not too small so that the joint probabilities can be calculated properly.
In the oscillating EF case, the net probability flux is also highly oscillating,
so it is convenient to study the integrated probability flux:
\begin{align}
  Q_{J,I} (t) = \int_0^t F_{J,I}(\tau)d \tau,
\end{align}
which presents the time averaged effect of the EF to the molecular conformation.

\subsection{Results and discussions}
\begin{figure}
  \centering
  \includegraphics[]{fig.ala/fig-meta-npt.eps}
  \caption{The probability of being in a metastable states under a constant EF.
    The warm-up time $t_{init} = 10$~ps. The red line stands for state $A_1$,
    green for $A_2$, dark blue for $B_1$, purple for $B_2$ and light blue
    for $C$.
  }
  \label{fig:tmp5}
\end{figure}
